---
title: "assignment27/28"
format:
  html:
    self-contained: true
---

## Setup and Packages

```{r setup}
library(sf)
library(terra)
library(osmdata)
library(dplyr)
library(ggplot2)
library(AOI)
```

## 1. Define River Object

```{r river-object}
aoi_fc <- AOI::geocode("Fort Collins, CO", bbox = TRUE)

iosm <- opq(bbox = st_bbox(aoi_fc)) %>%
  add_osm_feature(key = "waterway") %>%
  osmdata_sf()
osm_lines <- iosm$osm_lines

poudre_river <- osm_lines %>%
  filter(name == "Cache la Poudre River") %>%
  st_union() %>%
  st_as_sf()

river_length_raw <- st_length(poudre_river)
```

## 2. Define and Crop DEM Object

```{r dem-object}
Sys.setenv("AWS_NO_SIGN_REQUEST" = "TRUE")

dem_vrt <- rast("/vsis3/lynker-spatial/gridded-resources/dem.vrt")

dem_crs     <- crs(dem_vrt)
aoi_fc_proj <- st_transform(aoi_fc, dem_crs)
dem         <- crop(dem_vrt, vect(aoi_fc_proj))

names(dem) <- "elevation_cm"

dem
```

## 3. Reproject River and Extract Elevation

```{r extract-elev}
poudre_river_proj <- st_transform(poudre_river, dem_crs)
poudre_pts_proj  <- poudre_river_proj %>% st_cast("POINT")

elev_vals <- terra::extract(dem, vect(poudre_pts_proj), bind = FALSE)

profile <- poudre_pts_proj %>%
  bind_cols(elev_vals %>% select(elevation_cm)) %>%
  mutate(
    ID          = row_number(),
    elevation_m = elevation_cm / 100
  )

na_count <- sum(is.na(profile$elevation_cm))
message("Number of NA elevation values: ", na_count)
```

## 4. Compute Sinuosity

```{r sinuosity}
dist_straight_m <- as.numeric(
  st_distance(profile[1, ], profile[nrow(profile), ])
)

river_length_m <- as.numeric(st_length(poudre_river_proj))

sinuosity <- river_length_m / dist_straight_m
sinuosity
```

#This shows that the river is constantly twisting and turning and never stays straight for too long.  Yes this very much makes sense having been to the Poudre several times.

## 5. Compute Slope

```{r slope}
elev_change_m <- profile$elevation_m[nrow(profile)] - profile$elevation_m[1]

slope <- elev_change_m / river_length_m
slope
```

#197,781cm-146,730cm= 51,051cm.  The river is 203km long which makes the slope 0.0025.

## 6. Line Plot of Elevation Profile

```{r plot-profile-line}
ggplot(profile, aes(x = ID, y = elevation_m)) +
  geom_line(size = 1) +
  labs(
    x = "Point Index Along River",
    y = "Elevation (m)",
    title = "Cache la Poudre River Elevation Profile"
  ) +
  theme_minimal()
```

## 7. Spatial Map Colored by Elevation

```{r plot-profile-spatial}
ggplot() +
  geom_sf(data = profile, aes(color = elevation_m), size = 1) +
  scale_color_viridis_c(name = "Elevation (m)") +
  labs(
    title = "Cache la Poudre River Spatial Profile",
    subtitle = "Colored by Elevation"
  ) +
  theme_minimal()
```

